Reducing Acceptance Marks in Emerson-Lei Automata by QBF Solving

### 2023-03-20 paper/main.tex
changes in Fig.3

%\todo{Upravit experimenty: pridat dalsi 2 sady automatu.}
We have used the tool \texttt{genltl} of the Spot
library~\cite{duret.16.atva2} to generate LTL formulas from
literature. The parametrized formulas were generated for all
combinations of parameter values from 1 to 4. We used the tool
\texttt{ltlfilt} of the Spot library to simplify the formulas and
remove duplicates. Altogether, we have generated 348 unique LTL
formulas. These have been translated to deterministic TELA by two
state-of-the-art translators from the Owl
library~\cite{kretinsky.18.atva}, namely \texttt{ltl2dela} (known as
Delag)~\cite{muller.17.gandalf} and \texttt{ltl2dgra} (known as
Rabinzer~4)~\cite{kretinsky.18.cav}, and to nondeterministic TELA by
another two state-of-the-art translators, namely \texttt{ltl2tgba}
from the Spot library~\cite{duret.16.atva2} and
\texttt{ltl3tela}~\cite{major.19.atva}. Each of the four considered
tools failed on some formulas: it usually reached a timeout of 60
seconds or it produced an automaton that cannot be parsed by the Spot
library.
% produced by the individual tools contains less than 348 automata as not
% all of the formulas are successfully translated due to a timeout of 60
% seconds.
Further, we have removed automata with 0 or 1 acceptance marks as
there is a little point in reducing these. Table
\ref{tab:ltltranslators} shows the exact versions and homepages of
used translators. For each translator, it also provides the number of
fails, the number of produced automata with less than two marks, and
the number of automata with at least two marks. The last number is
typeset in bold as these automata are actually used for the
experimental evaluation of our reduction technique.

\begin{table}[t]
  \caption{LTL to automata translators used for the experimental
    evaluation and the numbers of fails and successfully constructed
    automata with less than two marks and with at least two marks.}
\label{tab:ltltranslators}
\centering
% \begin{tabular}{lr@{~~~~~~}lr}
% \toprule
% tool        & version&homepage & ~~~automata \\
% \midrule
% \texttt{ltl2dela}~~~ & 21.0 & \multirow{2}{*}{\url{https://owl.model.in.tum.de/}} & 129\\
% \texttt{ltl2dgra} & 21.0 & & 234\\
% \texttt{ltl2tgba} & 2.10.4 & \url{https://spot.lrde.epita.fr} & 70\\
% \texttt{ltl3tela} & 2.2.0  & \url{https://github.com/jurajmajor/ltl3tela} & 91\\
% \bottomrule
% \end{tabular}
%
\begin{tabular}{lrrrr}
  \toprule
  &&& \multicolumn{2}{c}{constructed automata with} \\ \cmidrule(lr){4-5}
  tool &\!\!\!version/homepage & fails & 0 or 1 mark & \textbf{at least 2 marks}\\
  \midrule
  \texttt{ltl2dela} & 21.0\tablefootnote{\label{fn:owl}\url{https://owl.model.in.tum.de/}}
       & 5 & 214 & \textbf{129} \\
  \texttt{ltl2dgra} & 21.0\footnotemark[1]
       & 12 & 102 & \textbf{234}\\
  \texttt{ltl2tgba} & 2.10.4\tablefootnote{\url{https://spot.lrde.epita.fr}}
       & 0 & 278 & \textbf{70}\\
  \texttt{ltl3tela} & 2.2.0\tablefootnote{\url{https://github.com/jurajmajor/ltl3tela}}
       & 18 & 239 & \textbf{91}\\
  \bottomrule
\end{tabular}
\end{table}

For each automata set (that is, the automata with at least two marks
produced by an individual LTL translator), we have applied all
single-level reductions and the multi-level reduction, always with
incremental approach and without reducing of the number of cubes. We
do not reduce the number of cubes as our primary aim is to reduce the
number of acceptance marks.
%\todo{Vysvetlit, proc neredukujeme pocet cubes?}
The timeout for each QBF query was set to 30 seconds.
%
All reductions have been performed by the tool \texttt{telatko}
built with Spot library version 2.10.4 and Z3 version 4.8.15. The
experiments have been run on a computer with Intel$^\text{®}$ Core™ i7-8700
processor and 32 GB of memory running Ubuntu 20.04.4. We used the tool
\texttt{autcross} of the Spot library to get the statistics of the
reduced automata and the running times.

@@

Figure~\ref{fig:quantile} presents the time spent by multi-level
reduction of automata in individual automata sets. The chart shows a
pleasing finding that for every set, most automata are reduced in
under a second and the high cumulative running times are caused by
a few complicated automata.

\todo{popisek Fig. 3 neni hezky. Je to wall time nebo CPU time? Je to
  kumulovany cas nebo pro kazdy automat zvlast? Zmenit popisky os v
  obrazku?}
\begin{figure}[b!]
  \caption{Performance of \telatko on different automata sets. The $x$
    axis represents the number of automata from each automata
    set that are processed by \telatko in the time shown on the $y$
    axis.}
\label{fig:quantile}
\bigskip
\centering
\input{quantile.tex}
\end{figure}

>>>>

%\todo{Upravit experimenty: pridat dalsi 2 sady automatu.}
We have used the tool \texttt{genltl} of the Spot
library~\cite{duret.16.atva2} to generate LTL formulas from
literature. The parametrized formulas were generated for all
combinations of parameter values from 1 to 4. We used the tool
\texttt{ltlfilt} of the Spot library to simplify the formulas and
remove duplicates. Altogether, we have generated 348 unique LTL
formulas. These have been translated to deterministic TELA by two
state-of-the-art translators from the Owl
library~\cite{kretinsky.18.atva}, namely \texttt{ltl2dela} (known as
Delag)~\cite{muller.17.gandalf} and \texttt{ltl2dgra} (known as
Rabinzer~4)~\cite{kretinsky.18.cav}, and to nondeterministic TELA by
another two state-of-the-art translators, namely \texttt{ltl2tgba}
from the Spot library~\cite{duret.16.atva2} and
\texttt{ltl3tela}~\cite{major.19.atva}. Three of the considered
translators failed on some formulas: it usually reached a timeout of
60 seconds or it produced an automaton that cannot be parsed by the
Spot library.
% produced by the individual tools contains less than 348 automata as not
% all of the formulas are successfully translated due to a timeout of 60
% seconds.
Further, we have removed automata with 0 or 1 acceptance marks as
there is a little point in reducing these. Table
\ref{tab:ltltranslators} shows the exact versions and homepages of
used translators. For each translator, it also provides the number of
fails, the number of produced automata with less than two marks, and
the number of automata with at least two marks. The last number is
typeset in bold as these automata are actually used for the
experimental evaluation of our reduction technique.

\begin{table}[t]
  \caption{LTL to automata translators used for the experimental
    evaluation and the numbers of fails and successfully constructed
    automata with less than two marks and with at least two marks.}
\label{tab:ltltranslators}
\centering
% \begin{tabular}{lr@{~~~~~~}lr}
% \toprule
% tool        & version&homepage & ~~~automata \\
% \midrule
% \texttt{ltl2dela}~~~ & 21.0 & \multirow{2}{*}{\url{https://owl.model.in.tum.de/}} & 129\\
% \texttt{ltl2dgra} & 21.0 & & 234\\
% \texttt{ltl2tgba} & 2.10.4 & \url{https://spot.lrde.epita.fr} & 70\\
% \texttt{ltl3tela} & 2.2.0  & \url{https://github.com/jurajmajor/ltl3tela} & 91\\
% \bottomrule
% \end{tabular}
%
\begin{tabular}{lrrrr}
  \toprule
  &&& \multicolumn{2}{c}{constructed automata with} \\ \cmidrule(lr){4-5}
  tool &\!\!\!version/homepage & ~~~~fails & less than 2 marks & \textbf{at least 2 marks}\\
  \midrule
  \texttt{ltl2dela} & 21.0\tablefootnote{\label{fn:owl}\url{https://owl.model.in.tum.de/}}
       & 5 & 214 & \textbf{129} \\
  \texttt{ltl2dgra} & 21.0\footnotemark[1]
       & 12 & 102 & \textbf{234}\\
  \texttt{ltl2tgba} & 2.10.4\tablefootnote{\url{https://spot.lrde.epita.fr}}
       & 0 & 278 & \textbf{70}\\
  \texttt{ltl3tela} & 2.2.0\tablefootnote{\url{https://github.com/jurajmajor/ltl3tela}}
       & 18 & 239 & \textbf{91}\\
  \bottomrule
\end{tabular}
\end{table}

To each automata set (that is, the automata with at least two marks
produced by an individual LTL translator), we have applied all
single-level reductions and the multi-level reduction, always with
incremental approach and without reducing of the number of cubes. We
do not reduce the number of cubes as our primary aim is to reduce the
number of acceptance marks.
%\todo{Vysvetlit, proc neredukujeme pocet cubes?}
The timeout for each QBF query was set to 30 seconds.
%
All reductions have been performed by the tool \texttt{telatko}
built with Spot library version 2.10.4 and Z3 version 4.8.15. The
experiments have been run on a computer with Intel$^\text{®}$ Core™ i7-8700
processor and 32 GB of memory running Ubuntu 20.04.4. We used the tool
\texttt{autcross} of the Spot library to get the statistics of the
reduced automata and the running times.

@@

Figure~\ref{fig:quantile} presents the time spent by multi-level
reduction on individual automata of each automata sets. The chart
shows a pleasing finding that for every set, most automata are reduced
in under a second and the high cumulative running times are caused by 
a relatively small number of complicated automata.

\begin{figure}[b!]
  \caption{Performance of \telatko on individual automata of each automata sets.
    The graph shows the time ($y$ axis) needed by \telatko
    to process the $x$-th automaton of the set, where automata of each set are ordered by their processing time.
    % The $x$ axis represents the number of automata from each automata
    % set that are processed by \telatko in the time shown on the $y$
    % axis.
  }
\label{fig:quantile}
\bigskip
\centering
\input{quantile.tex}
\end{figure}
